\section{Future Work}
At this point, we have a replacement class for vtkGPURayCastMapper that is more widely supported, faster, more easily extensible, and supports majority of the features of the old class. In the near future, our goal is to ensure that this mapper works as promised by integrating it into existing VTK applications such as ParaView and Slicer. Once these tasks are complete, we have some ideas on new features we would like to add to this mapper (outlined below). We would also like to solicit feedback from folks using the VTK volume mappers. What features do you need? Drop us a line at kitware@kitware.com, and let us know.

\subsection{2D Transfer Functions} Currently, volume rendering in VTK uses two 1D transfer functions, mapping scalar value to opacity and gradient magnitude to opacity. For some application areas, better rendering results can be obtained by using a 2D table that maps these two parameters into an opacity value. Part of the challenge in adding a new feature such as this to volume rendering in VTK is simply the number of volume mappers that have to be updated to handle it (either correctly rendering according to these new parameters or at least gracefully implementing an approximation). Once we have reduced the number of volume mappers in VTK, then adding new features such as this will become more manageable.

\subsection{Support for Depth Peeling} Currently, VTK correctly intermixes volumes with opaque geometry. For translucent geometry, you can obtain a correct image only if all translucent props can be sorted in depth order. Therefore, no translucent geometry can be inside a volume, as it would be, for example, when depicting a cut plane location with a 3D widget that represents the cutting plane as a translucent polygon.  Nor can a volume be contained within a translucent geometric object, as it would be if, for example, the outer skin of a CT data set was rendered as a polygonal isosurface with volume mappers used to render individual organs contained within the skin surface. We hope to extend the new vtkGPURayCastMapper to support the multipass depth peeling process, allowing for correctly rendered images with intersecting translucent objects.

\subsection{Overlapping Volumes}
It is currently possible to render overlapping volumes by taking advantage of the up to 4 independent components supported by the mapper (each component representing a different volume).  The limitation with this approach is that each of the overlapping volumes are required to be sampled in the same grid, hence all of the volumes are required to share the same dimensions.  Nonetheless, in order to extend the mapper to support overlapping volumes sampled in grids with different dimensions, rays should be casted through proxy geometry bounding the N overlapping volumes to be rendered and separately sampling and compositing their fetched texture values in the fragment shader.
Improved Rendering of Labeled Data: Currently, VTK supports binary masks and only a couple of very specific versions of label mapping. We know that our community needs more extensive label mapping functionality – especially for medical datasets. Labeled data requires careful attention to the interpolation method used for various parameters. (You may wish to use linear interpolation for the scalar value to look up opacity, but, perhaps, select the nearest label to look up the color.) We plan to solicit feedback from the VTK community to understand the sources of labeled data and the application requirements for visualization of this data. We then hope to implement more comprehensive labeled data volume rendering for both the CPU and GPU mappers.