\section{Background}
\label{relatedwork}

Volume visualization provides a two-dimensional view of three-dimensional
uniformly sampled data by sampling and projecting the data onto a 2D
projection plane.  While many techniques exist for volume rendering such as
marching cubes~\citep{lorensen_marching_1987}, image
splatting~\citep{westover_footprint_1990}, texture
slicing~\citep{rezk-salama_interactive_2000, engel_high-quality_2001} and ray
casting~\citep{hsu_segmented_1993, ma_parallel_1995, ma_scalable_1997,
heng_gpu-based_2005}, ray casting has become pervasive on modern graphics
hardware. Ray casting techniques provides a higher quality of rendering and
various options to optimize the rendering performance such as early ray
termination and space leaping~\citep{yagel_accelerating_1993}.

While ray casting is a well-known technique and various implementations exist,
providing high-quality volume visualization for datasets of adequate
resolution, supporting a variety of spacing and formats, with or without
intermixed geometry rendering, on multiple platforms (desktop, Cave, Head
Mounted Displays (HMD)) and across operating systems (Apple, Linux, Window), at
interactive speeds is still a challenge.

Few open source software systems exist for volume visualization, such as
Voreen~\citep{meyer-spradow_voreen:_2009} which initiated at Department of
Computer Science at the University of M\"unster, Germany in 2004.  It provides
features such as isosurface rendering, maximum intensity projection
(MIP)~\citep{wallis_three-dimensional_1989}, support for 1D and 2D transfer
functions, and support for different illumination models (Phong, tone, ambient
occlusion). Both Voreen and VTK use data-flow networks. However, Voreen is a
volume visualization library whereas VTK is a scientific visualization library
and provides better support for geometry and gridded datasets. While Voreen
provides a rapid prototyping environment, VTK volume visualization aims for
production quality, performance, and ubiquitousness. Additionally, Voreen only
supports high-end desktop devices whereas VTK supports multiple platforms
natively, bridging the gap between academic research and practical application.
Other volume visualization capable tools such as ParaView, Slicer, and MayaVi,
all are using VTK underneath as the underlying visualization library and therefore
will benefit from our work. ParaView has been switched to use our work and
updating the slicer to use our mapper is in progress.

Another volume rendering capable tool Amira~citep{stalling2005amira} provides
multiple modules for volume visualization namely: voltex, volren, and volume
rendering module. In the context of our work, the volume rendering module of
Amira is most relevant. It provides support for GPU rendering, iso-surface, MIP,
and multi-volume support. In comparison to our work, it provides limited
support for data larger than main memory as it needs conversion to LDA format,
lack of other volume rendering modes such as MinIP, off-screen rendering,
and gradient-opacity and is closed-source. OpenWalnut is another open-source
tools that for visualization of medical and bio-signal data but provides very
basic volume visualization support but does offer off-screen rendering support.
Another example of an open source volume rendering software system is
ImageVis3D~\citep{cibc_imagevis3D:_2016}, developed by the researchers at the
University of Utah. ImageVis3D is an application as opposed to a library and
provides support for large volume data, desktop as well as mobile devices, MIP,
1D and 2D transfer functions amongst many others.  It does not support Virtual
Reality environments and the flexibility of data-flow networks. Other data
type specific volume rendering libraries such as
Voxx~\citep{clendenon_voxx:_2002} and
ClearVolume~\citep{royer_clearvolume:_2015} provide visualization of
biological and light-sheet microscopy.
PyMOL~\citep{schrodinger_llc_pymol_2015} provides volume visualization
capabilities for molecular datasets only. Finally, hardware architecture
specific volume visualization libraries such as
OSPRay~\citep{wald_ospray_2017} and NVIDIA\textsuperscript{\textregistered}
Optix\textsuperscript{\texttrademark}~\citep{parker_optix:_2010} provide fast
large volume data visualization capabilities on Intel CPU and Nvidia GPU's but
lack support for non-native hardware.

Many volume rendering APIs exist apart from Voreen and ImageVis3D. However,
except for Voreen, advanced volume visualization research performed by the
industry and academic communities are not publicly available. By providing an
open source volume rendering engine that supports multiple platforms natively,
we hope to bridge the gap between academic research and industrial applications
in the field of volume visualization.
